MLH1 Results Outline

### 2020-02-24 doc/Results.rmd
small editing changes

total SC t tests
 p = `r Q2tot.sc_t.test_High.Low$p.value`
**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

musc and mol ttests
mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` Mole have significant difference between mouse means for total sc, between the high and low srains

musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`
the total sc mouse means are not significant for the high and low strains in musc

Additionally we ran similar tests using the short and long bivalent data sets which help to alimeorate potential chromosome size effects.

@@

mol-short

mol-long
very few significant comparisons>

@@

across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).

@@

(the subspecies effect is tested in M1)

@@

strain effects from M3; G, LEW, MSM, SKIVE and TOM are significant

strain effects  (basically all strain effects are significant) (tech effects / error?)
All strain effects were significant when tested in M2 (Supplemental table X)

In the short bivalent data set -- analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`). And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).

For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW
G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.

@@

24sep14_PWD_m1 is lower than the others
5jul17_MSM_m1 is the low MSM

The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)

@@

The mouse averages for IFDs were significantly longer in the high rec groups ( p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^). The pattern is confirmed with t.test comparing groups within the same subepcies. **IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.**

@@

For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs). Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^. The Mol\*chrmLength is negative (IN MOL as chrm length got longer - IFD gets smaller -- also in SKIVE\*chrm

For the strain only model,

@@

The general chromsome Length effect is that

The single fixed effects can be interperted as, with each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF). All of the high rec strains and MOLF have significantly longer IFD^PER^.

There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.

@@

<None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.>

Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent. This also indicated that interference distance may evolved in the house mouse complex.

@@

<below chunk goes with the above chunk>
```{r long.biv.plots, echo=FALSE, eval=FALSE}

>>>>

**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

Molossinus have significant difference between mouse means for total sc, between the high and low strains (mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`). The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).

<Additionally we ran similar tests using the short and long bivalent data sets which help to alimeorate potential chromosome size effects.>

@@

mol-short
mol-long
very few significant effects>

@@

<across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).>

@@

@@

In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`). 
<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>

<For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW, G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.>

@@

<24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM>

<The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)>

@@

The mouse averages for IFDs were significantly longer in the high rec groups ( p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^). The pattern is confirmed with t.test comparing groups within the same subepcies.

@@

For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs). Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^. The Mol\*chrmLength is negative

<IN MOL as chrm length got longer - IFD gets smaller -- also in SKIVE x chrm>

For the strain only model (M2),

@@

With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF). All of the high rec strains and MOLF have significantly longer IFD^PER^. There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.

@@

 **IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.**

@@

<HAND measure METRICS, below chunk goes with the above chunk>
```{r long.biv.plots, echo=FALSE, eval=FALSE}
